\vspace{-0.25em}\thetitle

### 2009-07-24 paper.tex
...

Our answer to the question of why low-dimensional linear structure appears in the feature-frequency representation of network traffic can be expressed as a stochastic model of flow behavior.
% A mixture model, to be specific.
We propose viewing each flow as having a probability distribution of possible behaviors it could exhibit.
% We propose that the behavior of each flow is drawn from a probability distribution on a space possible behaviors.
We hypothesize that the behavior distribution for most flows is a mixture of a relatively small set of ``basic behaviors.''
Moreover, only even smaller subsets of these basic behaviors are typically combined with each other.
Under these assumptions, we can express the distribution of each flow's behaviors as a finite mixture model~\cite{McLachlan00}:
\begin{align}\eqn{mixture-model}
  q_i(x) = \sum_{j=1}^r w_{ij} p_j(x),
\end{align}
Here $q_i$ and $p_j$ are probability density functions, and $w_{ij}$ are nonnegative weights, summing to unity for each $i$.
\Equation{mixture-model} is expressed succinctly as matrix multiplication.
Writing $Q_{ik} = q_i(k)$, $W_{ij} = w_{ij}$, and $P_{jk} = p_j(k)$, we have:
\begin{align}\eqn{mixture-model-matrix}
  Q = WP.
\end{align}
The number of basic behaviors, $r$, is the maximum possible rank of the feature distribution matrix, $Q$.
Next, partition the rows of $P$ into small sets which can have simultaneously non-zero weights: $P = [P_1;\cdots;P_g]$ using Matlab concatenation notation;
assume the rows of $P$ are ordered to allow this.
We can also assume that the rows of $Q$ and $W$ are ordered so that we can write $W = W_1 \directsum \cdots \directsum W_g$.

% Mixture modeling expresses probability densities as convex combinations of constituent probability distributions:
% \begin{align}\eqn{mixture-model}
%   q_i(x) = \sum_{j=1}^r w_{ij} p_j(x),
% \end{align}
% Here $q_i$ and $p_j$ are density functions, and $w_{ij}$ are nonnegative weights, summing to unity for each $i$.
% \Equation{mixture-model} says that each of the distributions $q_i$ is a convex mixture of the

% The question of why we see low-dimensional linear structure in the feature-frequency representation of network traffic is deeply connected to the nature of such linear representations themselves.
% But the answer to the question also requires us to posit a hypothesis about the nature of network traffic.

>>>>

Our answer to the question of why low-dimensional linear structure appears in the feature-frequency representation of network traffic can be expressed as a stochastic model of flow behavior.
% A mixture model, to be specific.
We propose viewing each flow as having a probability distribution of possible behaviors it could exhibit.
% We propose that the behavior of each flow is drawn from a probability distribution on a space possible behaviors.
We hypothesize that the behavior distribution for most flows is a mixture of a relatively small set of ``basic behaviors.''
Moreover, only even smaller subsets of these basic behaviors are typically combined with each other.
Under these assumptions, we can express the distribution of each flow's behaviors as a finite mixture model~\cite{McLachlan00}:
\begin{align}\eqn{mixture-model}
  q_i(x) = \sum_{j=1}^r w_{ij} p_j(x),
\end{align}
Here $q_i$ and $p_j$ are probability density functions, and $w_{ij}$ are nonnegative weights, summing to unity for each $i$.
\Equation{mixture-model} is expressed succinctly as matrix multiplication.
Writing $Q_{ik} = q_i(k)$, $W_{ij} = w_{ij}$, and $P_{jk} = p_j(k)$, we have:
\begin{align}\eqn{mixture-model-matrix}
  Q = WP.
\end{align}
The number of basic behaviors, $r$, is the maximum possible rank of the feature distribution matrix, $Q$.
Moreover, we can partition the rows of $P$ into disjoint classes such that $w_{ij_1}$ and $w_{ij_2}$ are both non-zero only if $j_1$ and $j_2$ are in the same class.
Thus, each row of $Q$ is associated with exactly one class, and all the points associated with a class lie in the subspace spanned by its rows in $P$.

This model is an ideal, and observed behaviors have noise introduced by sampling, imperfect adherence to the model, and outliers.
Yet the properties of the model explain the structures seen in \Figure{svd}.
However, this is only one hypothesis that could fit the data.
Like any hypothesis, it must be tested.
The rest of this paper, aside from providing a useful, practical application, serves as a hypothesis test:
we try to recover the matrices $W$ and $P$ from our imperfect observations of $Q$ and use the recovered model to predict flow behaviors.
If the recovered model makes accurate predictions, this provides evidence that our hypothesis is correct.
